|

### 2023-09-07 chapter_4/src/prr/dcr_color_and_non-political_behaviors_prr.qmd
a mockup of the treatments
a little bit more editing

The implications of such findings corroborate existing work that suggests that there are an increasing number of cultural signals that people use to guess the partisanship of not just other individuals but groups of people, and that as a result partisanship and polarization are bleeding into our non-political decision-making. The implications of the argument advanced in the chapter build upon these existing findings by providing one causal mechanism as to the connection between cultural cues conveyed visually and the cognitive processing of such information. The chapter not only generalizes findings from the previous chapters to groups of people and to non-political attitudes and behaviors, but it contributes to a literature suggesting that political polarization is pervasive and is approaching universal.

@@

While political polarization certainly reflects a growing gap and consolidation of two camps of policy preferences and attitudes [@dellaposta_pluralistic_2020;@luders_attitude_2023], it manifests as something more consequential for our every-day experiences. Political polarization has come to reflect social groups that distinguish "us" versus "them" with increasingly fewer and fewer non-political social groups that are not wrapped up in our partisan group [@mason_uncivil_2018]. That is, we are not only holding fewer and fewer cross-cutting attitudes that moderate our political viewpoints [@luders_attitude_2023], but we also have fewer and fewer cross-cutting social groups that encourage interaction and experiences with perspectives of out-partisans [@mason_uncivil_2018]. The consequence of such phenomena is that we are becoming increasingly willing to harbor strong negative feelings towards those of the other political party and strong positive affectations for members of our political party [@iyengar_affect_2012;@iyengar_fear_2015].

@@

Building on this set of evidence, a growing area of the literature suggests that we are able to make inferences about the partisanship about other individuals, but we are able to do this for groups of people through the increasing politicultural distinctions between Republicans and Democrats. @hetherington_prius_2018 demonstrates that Republicans and Democrats are not just different in their policy preferences but rather that they live in different places and have different lifestyles. Others have replicated and built upon this work by suggesting that while there are many different types of lifestyles that Republicans and Democrats can take on, there is increasingly clear distinction between Republicans and Democrat archetypes [@rogers_politicultural_2022]. These politicultural distinctions are clear enough that we train a computer to make accurate guesses (upwards of 80%) of the partisanship by using only images of the cars in a particular neighborhood [@gebru_using_2017].

@@

As discussed thus far in the chapter, non-political objects are often associated with particular partisan leanings. For example, the modal car in a neighborhood conveys the partisan composition of the whole neighborhood [@hetherington_prius_2018;@gebru_using_2017]. In the same spirit as the rest of this book, we do not need to necessarily distinguish between whether someone drives a hybrid sedan and a Ford pickup truck. Information giving insight into one's partisanship may even be as subtle as their color choice in car. For a reader who thinks "hypothesizing about car color choice and partisanship is a reach," I agree. As stated at the outset of the chapter, the goal of the chapter is to extend past applications of the snap-judgment model to explicitly partisan objects such as yard signs as we did in Chapter 2 and to really challenge the model by looking at non-political objects that many never associate with partisanship.

Recall the snap-judgment model as depicted in @fig-theory. The model suggests that so long as politics is salient, the colors red and blue should act as a symbolic association with Republicans and Democrats, respectively. This association will trigger a pre-conscious and potent process that involves an activation of an affective state depending on one's prior experiences with Republicans and Democrats. If, when viewing those colors, activates memories of Republicans and Democrats that are negative, then your affective state will have a negative valance. If those activated memories are positive, then your affective state will have a positive valance. You will associate your affective state with the object that started the process due to the object's association with memories and their associations to Republicans and Democrats. This snap-judgment that you form will not only influence your expressed attitudes toward that object, but will influence motivations to engage more with the object or to avoid it. With new information, we can refine which memories we use to evaluate the object and can therefore form a response that is better informed. However, the snap-judgment persists and will have influences on which memories connected to that new information we are accessing. Therefore, the snap-judgment still has effects on our attitude expression and behaviors even with new information.

@@

To test these hypotheses, I conduct an experiment involving a fictional real estate posting. The treatment involves randomly assigning participants to view one of two versions of a fictional real estate posting. The characteristics of the neighborhood are exactly the same between the two conditions except for the "favorite car" in the neighborhood. In the first condition the "favorite car" is a red SUV. In the second condition, the "favorite car" is the same make and model of SUV as in the first condition, but blue. Participants are then given a scale from 0-10 to rate the neighborhood and are presented with a Likert scale to indicate, "How much would [they] like to live in this neighborhood?"

@@

>>>>

The implications of such findings corroborate existing work that suggests that there are an increasing number of cultural signals that people use to infer the partisanship of not just other individuals but groups of people, and consequently partisanship and polarization are bleeding into our non-political decision-making. The implications of the argument advanced in the chapter build upon these existing findings by providing one causal mechanism as to the connection between cultural cues conveyed visually and the cognitive processing of such information. The chapter not only generalizes findings from the previous chapters to groups of people and to non-political attitudes and behaviors, but it contributes to a literature suggesting that political polarization is pervasive as a basis of forming attitudes and is approaching universal.

@@

While political polarization certainly reflects a growing gap and consolidation of two groups of policy preferences and attitudes [@dellaposta_pluralistic_2020;@luders_attitude_2023], it manifests as something more consequential for our every-day experiences. Political polarization has come to reflect social groups that distinguish "us" versus "them" with increasingly fewer non-political social groups that are not wrapped up in our partisan group [@mason_uncivil_2018]. That is, we are not only holding fewer cross-cutting attitudes that moderate our political viewpoints [@luders_attitude_2023], but we also have fewer cross-cutting social groups that we identify with to encourage interaction and experiences with perspectives of out-partisans [@mason_uncivil_2018]. The consequence of such phenomena is that we are becoming increasingly willing to harbor strong negative feelings towards those of the other political party and strong positive affectations for members of our political party [@iyengar_affect_2012;@iyengar_fear_2015].

@@

Building on this set of evidence, a growing area of the literature suggests that we are not just able to make inferences about the partisanship about other individuals, but we are also able to do this for groups of people through the increasing politicultural distinctions between Republicans and Democrats. @hetherington_prius_2018 demonstrates that Republicans and Democrats are not just different in their policy preferences but rather that they live in different places and have different lifestyles. Others have replicated and built upon this work by suggesting that while there are many different types of lifestyles that Republicans and Democrats can take on, there is increasingly clear distinction between Republicans and Democrat archetypes [@rogers_politicultural_2022]. These politicultural distinctions are clear enough that we train a computer to make accurate guesses (upwards of 80%) of the partisanship by using only images of the cars in a particular neighborhood [@gebru_using_2017].

@@

As discussed thus far in the chapter, non-political objects are often associated with particular partisan leanings. For example, the modal car in a neighborhood conveys the partisan composition of the whole neighborhood [@hetherington_prius_2018;@gebru_using_2017]. In the same spirit as the rest of this book, we do not need to necessarily distinguish between complex forms of information such as whether someone drives a hybrid sedan and a Ford pickup truck. Information giving insight into one's partisanship may even be as subtle as their color choice in car. For a reader who thinks "hypothesizing about car color choice and partisanship is a reach," I agree. As stated at the outset of the chapter, the goal of the chapter is to extend past applications of the snap-judgment model to explicitly partisan objects such as yard signs as we did in Chapter 2 and to really challenge the model by looking at non-political objects that we do not readily associate with partisanship.

Recall the snap-judgment model as depicted in @fig-theory. The model suggests that so long as politics is salient, the colors red and blue should act as a symbolic association with Republicans and Democrats, respectively. This association will trigger a pre-conscious and potent process that involves an activation of an affective state depending on one's prior experiences with Republicans and Democrats. If, when viewing those colors, activated memories of Republicans and Democrats that are negative, then your affective state will have a negative valance. If those activated memories are positive, then your affective state will have a positive valance. You will associate your affective state with the object that started the process due to the object's association with memories and their associations to Republicans and Democrats. This snap-judgment that you form will not only influence your expressed attitudes toward that object, but will influence motivations to engage more with the object or to avoid it. With new information, we can refine which memories we use to evaluate the object and can therefore form a response that is better informed. However, the snap-judgment persists and will have influences on which memories connected to that new information we are accessing. Therefore, the snap-judgment still has effects on our attitude expression and behaviors even with new information.

@@

To test these hypotheses, I conduct an experiment involving a fictional real estate posting. The treatment involves randomly assigning participants to view one of two versions of a fictional real estate posting. The characteristics of the neighborhood are exactly the same between the two conditions except for the "favorite color of car" in the neighborhood. In the first condition the "favorite color of car" is reported by displaying a red swatch. In the second condition, the "favorite color of car" reported by displaying a blue swatch. Participants are then given a scale from 0-10 to rate the neighborhood and are presented with a Likert scale to indicate, "How much would [they] like to live in this neighborhood?"

::: {#fig-treatments layout-ncol=2}

![Blue treatment](./assets/treatments/blue_mockup.png){#fig-blue-treatment}

![Red treatment](./assets/treatments/red_mockup.png){#fig-red-treatment}

Treatments
:::

@@

## Modeling strategy

## Simulations
